\bf Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-16 manuscript/main.tex
proposition 1 proof anchor

\begin{equation}\label{def:squared_empirical_cov}
	 	\widehat{\textnormal{Cov}} 
	 	( \langle 
	 	\widetilde{W},
	 	\xi \rangle_{\mathcal{H}}, Y )
	 	:=
	 	\frac{1}{n} 
	 	\sum_{i=1}^n
	 	y_i
	 	\langle \widetilde{W}_i, \xi \rangle_{\mathcal{H}}
	 	.
	 \end{equation}
We denote the estimated PLS component direction as
	\begin{equation}\label{def:maximizer_squared_empirical_cov}
		\hat{\xi} 
= 
(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\xi}}_{K+1}
)
:= \arg \max_{\xi \in \mathcal{B}^K \times \mathbb{R}^p,
\| \xi  \|_\mathcal{H} = 1
 } \widehat{\textnormal{Cov}}^2 
\bigl(
\langle \widetilde{W}, \xi \rangle_{\mathcal{H}}, Y 
\bigr),
	\end{equation}
%
	where each functional part $\hat{\xi}_j $  is expanded as
$
		\hat{\xi}_j(t) := 
		\sum_{m=1}^M \hat{\gamma}_{jm} \, b_m(t),
$
and $\hat{\boldsymbol{\xi}}_{K+1}$ is the scalar part. This maximizer can also be represented as a Euclidean vector
\begin{equation}\label{def:maximizer_euclidean_form}
	( \hat{\boldsymbol{\gamma}}_1^\top, \ldots, \hat{\boldsymbol{\gamma}}_K^\top, \hat{\boldsymbol{\xi}}_{K+1}^\top )^\top
	\in \mathbb{R}^{MK+p},
\end{equation}
where the coefficient vectors $\hat{\boldsymbol{\gamma}}_1, \ldots, \hat{\boldsymbol{\gamma}}_K$ are defined as in \eqref{def:coef_vector_form}.
Proposition \ref{proposition:eigen_noregul} reformulates  the problem \eqref{def:maximizer_squared_empirical_cov} as  a generalized Rayleigh quotient.
%
\begin{proposition}\label{proposition:eigen_noregul}
Let
$
	(\mathbb{B}, \mathbb{B}^{\prime \prime}, \Theta, \mathbf{y})
$
denote the observed data defined in \eqref{def:problem_instance}.
At the $l$-th iteration of the PLS algorithm, the unit-norm maximizer of  the squared empirical covariance defined in \eqref{def:maximizer_squared_empirical_cov}, represented as the Euclidean vector form as in \eqref{def:maximizer_euclidean_form},
is obtained by solving 
\begin{equation}	\label{eq: Regularized generalized rayleigh quotient equation}
	\left(
	\hat{\boldsymbol{\gamma}}_1^\top, \ldots, \hat{\boldsymbol{\gamma}}_K^\top, \hat{\boldsymbol{\xi}}_{K+1}^\top
	\right)^\top
	=
	\arg \max_{\boldsymbol{\xi} \in \mathbb{R}^{MK+p}}   \boldsymbol{\xi}^\top V \boldsymbol{\xi}
	\quad \text{subject to} \quad \boldsymbol{\xi}^\top \mathbb{B} \boldsymbol{\xi} = 1.
\end{equation}
where
\begin{equation}\label{def:V_matrix}
V := \frac{1}{n^2} (\mathbb{B} \Theta^\top \mathbf{y})(\mathbb{B} \Theta^\top \mathbf{y})^\top \in \mathbb{R}^{(MK+p) \times (MK+p)}.
\end{equation}
\end{proposition}
Since $V$ is a rank 1, the solution  is  obtained by matrix multiplication and normalization:
	\begin{proposition} \label{proposition:linear_noregul}
	We can compute  $\hat{\boldsymbol{\gamma}}_1, \ldots, \hat{\boldsymbol{\gamma}}_K$ and $\hat{\boldsymbol{\xi}}_{K+1}$ in Proposition \ref{proposition:eigen_noregul} as 
\begin{equation*}
	\hat{\boldsymbol{\gamma}}_j  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	\Theta_j^\top \mathbf{y},
	\quad
	%%%%%%%%%%%%%%%%
	\hat{\boldsymbol{\xi}}  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\mathbf{y}^\top
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			+ 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	Z^\top \mathbf{y}.
\end{equation*}
\end{proposition}
The expressions above represent matrix–vector multiplications normalized by a common factor. 
Proofs of Propositions \ref{proposition:eigen_noregul} and \ref{proposition:linear_noregul} are provided in Appendix~\ref{section:proof:proposition:linear_noregul_system_unregularized}.
Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.

>>>>

\begin{equation}\label{def:squared_empirical_cov}
	 	\widehat{\textnormal{Cov}} 
	 	( \langle 
	 	\widetilde{W},
	 	\xi \rangle_{\mathcal{H}}, Y )
	 	:=
	 	\sum_{i=1}^n
	 	\langle \widetilde{W}_i, \xi \rangle_{\mathcal{H}}
	 	.
We denote the estimated PLS component direction as
	\begin{equation}\label{def:maximizer_squared_empirical_cov}
		\hat{\xi} 
(\hat{\xi}_1, \ldots, \hat{\xi}_K, \hat{\boldsymbol{\xi}}_{K+1}
:= \arg \max_{\xi \in \mathcal{B}^K \times \mathbb{R}^p,
\| \xi  \|_\mathcal{H} = 1
 } \widehat{\textnormal{Cov}}^2 
\bigl(
\bigr),
	where each functional part $\hat{\xi}_j $  is expanded as
		\hat{\xi}_j(t) := 
		\sum_{m=1}^M \hat{\gamma}_{jm} \, b_m(t),
and $\hat{\boldsymbol{\xi}}_{K+1}$ is the scalar part. This maximizer can also be represented as a Euclidean vector
\begin{equation}\label{def:maximizer_euclidean_form}
	( \hat{\boldsymbol{\gamma}}_1^\top, \ldots, \hat{\boldsymbol{\gamma}}_K^\top, \hat{\boldsymbol{\xi}}_{K+1}^\top )^\top
	\in \mathbb{R}^{MK+p},
where the coefficient vectors $\hat{\boldsymbol{\gamma}}_1, \ldots, \hat{\boldsymbol{\gamma}}_K$ are defined as in \eqref{def:coef_vector_form}.
Proposition \ref{proposition:eigen_noregul} reformulates  the problem \eqref{def:maximizer_squared_empirical_cov} as  a generalized Rayleigh quotient.
\begin{proposition}\label{proposition:eigen_noregul}
Let
	(\mathbb{B}, \mathbb{B}^{\prime \prime}, \Theta, \mathbf{y})
denote the observed data defined in \eqref{def:problem_instance}.
At the $l$-th iteration of the PLS algorithm, the unit-norm maximizer of  the squared empirical covariance defined in \eqref{def:maximizer_squared_empirical_cov}, represented as the Euclidean vector form as in \eqref{def:maximizer_euclidean_form},
is obtained by solving 
\begin{equation}	\label{eq: Regularized generalized rayleigh quotient equation}
	\left(
	\hat{\boldsymbol{\gamma}}_1^\top, \ldots, \hat{\boldsymbol{\gamma}}_K^\top, \hat{\boldsymbol{\xi}}_{K+1}^\top
	\right)^\top
	\arg \max_{\boldsymbol{\xi} \in \mathbb{R}^{MK+p}}   \boldsymbol{\xi}^\top V \boldsymbol{\xi}
	\quad \text{subject to} \quad \boldsymbol{\xi}^\top \mathbb{B} \boldsymbol{\xi} = 1.
where
\begin{equation}\label{def:V_matrix}
V := \frac{1}{n^2} (\mathbb{B} \Theta^\top \mathbf{y})(\mathbb{B} \Theta^\top \mathbf{y})^\top \in \mathbb{R}^{(MK+p) \times (MK+p)}.
The proof of Proposition \ref{proposition:eigen_noregul}
is provided in Appendix \ref{section:proof:proposition:linear_noregul_system_unregularized}.
Since $V$ is a rank 1, the solution  is  obtained by matrix multiplication and normalization:
	\begin{proposition} \label{proposition:linear_noregul}
	We can compute  $\hat{\boldsymbol{\gamma}}_1, \ldots, \hat{\boldsymbol{\gamma}}_K$ and $\hat{\boldsymbol{\xi}}_{K+1}$ in Proposition \ref{proposition:eigen_noregul} as 
	\hat{\boldsymbol{\gamma}}_j  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	\Theta_j^\top \mathbf{y},
	\quad
	%%%%%%%%%%%%%%%%
	\hat{\boldsymbol{\xi}}  = \frac{1}{\sqrt{\sum_{j=1}^K 
			\Theta_j
			B \Theta_j^\top \mathbf{y} 
			\mathbf{y}^\top Z Z^\top \mathbf{y}}}
	Z^\top \mathbf{y}.
The expressions above represent matrix–vector multiplications normalized by a common factor. 
Proofs of Propositions \ref{proposition:eigen_noregul} and \ref{proposition:linear_noregul} are provided in Appendix~\ref{section:proof:proposition:linear_noregul_system_unregularized}.
Although the unnormalized coefficients are computed separately for the functional and scalar part of the PLS component direction, the normalization step couples them, allowing the components to influence each other. This enables the procedure to account for the correlation between functional and scalar components.
